**Accuracy and precision of 16S rRNA gene sequence assignments to OTUs is more important than the stability of OTUs**

### 2015-10-02 Schloss_Cluster_PeerJ_2015.Rmd
Editing
Add commas between numbers
Writing and editing

options(scipen=999)
```

@@

It is worth noting that the entire design of the He study is artificial. First, their analysis was based on a single sample. Researchers generally have dozens or hundreds of samples that are pooled and clustered together to enable comparison across samples. Second, all of the sequence data from these datasets is pooled for a single analysis. No one would ever perform an analysis based on a subset of their data. Because of these points, the value of identifying stable OTUs is unclear.

@@

@@

## Conclusions
* Stability is meaningless in context of accuracy and precision of clustering assignments
* *De novo* clustering algorithms outperform reference-based approaches.
* The best algorithm is data-specific and researchers should evaluate the clustering obtained by multiple methods

* Don't use USEARCH to cluster sequences
* VSEARCH is an excellent alternative

* Reference-based clustering has many problems

@@

>>>>

options(scipen=999)

inline_hook <- function(x){
  if(is.numeric(x)){
    paste(format(x,digits=2))
  }
}
knitr::knit_hooks$set(big.sep=',')
```
\newpage

## Abstract

\newpage

@@

@@

* Able to parallelize

@@

## Conclusions
It is worth noting that the entire design of the He study was artificial. First, their analysis was based on a single soil sample. Researchers generally have dozens or hundreds of samples that are pooled and clustered together to enable comparison across samples. Second, all of the sequence data from these datasets is pooled for a single analysis. It is unclear why anyone would ever perform an analysis based on a subset of their data (REFS). Because of these points, the value of identifying stable OTUs is unclear. Greater emphasis should be placed on obtaining an optimal balance between splitting similar sequences into separate OTUs and merging disparate sequences into the same OTU. Through the use of the pairwise distances between sequences, we were able to use the MCC to demonstrate that, in general, the AL, AGC, and DGC methods perform better than the others. Although there is concern that running the algorithms multiple times yields different clusterings, we have shown that there is little variation in their MCC values. This suggests that the different clusterings by the same algorithm are equally good. Finally, it is impossible to obtain a clustering with no false positives or false negatives and the optimal algorithm may vary by dataset. With this in mind, researchers are encouraged to run the AL and VSEARCH DGC algorithms and calculate and report their MCC values. MCC values can be calculated in the mothur software package using the `sens.spec` command.

Our analysis of those algorithms that implemented USEARCH as a method for clustering sequences revealed that its heuristic limited its sensitivity. When we replaced USEARCH with VSEARCH, the clustering quality improved. Although there may be parameters in USEARCH that can be tuned to improve the heuristic, these parameters are likely dataset dependent. Based on the data presented in this study, its availability as an open source, and free program, VSEARCH should replace USEARCH in these clustering algorithms. Furthermore, although not tested in our study, VSEARCH can be parallelized leading to potentially significant improvements in speed. Although USEARCh and VSEARCH do utilize aligned sequences, it is important to note that a sequence curation pipeline including denoising, alignment, trimming to a consistent region of the 16S rRNA gene, and chimera checking are critical to making proper inferences (REF).

* Reference-based clustering has many problems
* Open-reference clustering issues
* Implemented in numerous pipelines for large projects

@@

\newpage

## References
